Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2024-06-30 lqg-strings.tex
removed affiliation, added to discussion section, minor edit section 2

\author{Deepak Vaid}
\email{dvaid79@gmail.com}
\affiliation{National Institute of Technology, Karnataka, India}
\author{Luigi Teixeira de Sousa}
\email{luigi.tiraque@gmail.com}
\affiliation{Universidade Federal de São Carlos, Brasil}

@@

	The second is an action in which the background geometry is described in terms of the pullback of the connection which describes the bulk geometry to the worldsheet. The resulting action is in the form of a gauged sigma model, where the spacetime co-ordinates are now SO(D,1) vectors. We find that for the particular case of a constant background connection the action reduces to the bimetric action discussed above. We discuss classical solutions and quantization strategies for this action and its implications for the broader program of unifying string theory and loop quantum gravity.
\end{abstract}

@@

In a very real sense string theory is a quantum theory of geometry. After all, the Nambu-Goto action:
\begin{equation}\label{eqn:ng-action}
	S_{NG} = -T \int \diff ^2 x \sqrt{- \det h}
\end{equation}
is nothing more than the \emph{area} of the string worldsheet. So when we are constructing the quantum theory of the string we are constructing a theory in which the fundamental excitations are geometric in nature. What we would like to do is to modify the string action in a way which incorporates the insight from LQG that at the Planck scale there is a minimum quantum of area. One simple way in which to do this is to modify the action \eqref{eqn:ng-action} as follows:
\begin{equation} \label{eqn:NG-area-corrected}
	S_{M N G} = - T \int \diff ^2 x \sqrt{- (h + k \Delta)}.
\end{equation}
Here, $k = l_{pl} / l_s$, where $l_{pl}$ is the Planck scale, the scale at which quantum geometric effects become important; $l_s$ is the string scale where quantum geometric effects due to area quantization become small; $\Delta$ is a scalar density of weight 2 which can be taken to be the determinant of some worldsheet tensor $\Delta _{a b} (x)$ and corresponds to the square of the quantum of area at the point $x$ where the spin foam pierces the WS. In the limit that $l_s \gg l_{pl}$, $k \ll 1$ and the string action reduces to the conventional Nambu-Goto action with small corrections coming from quantum geometry. In this limit we can expand the integrand as follows:
\begin{equation} \label{eqn:inverse-metric-correction}
	\sqrt{- (h + k \Delta)} \sim \sqrt{- h}\left(1 - \frac{1}{2} \frac{k \Delta}{(- h)} + \mc{O}(k^2/h^2)\right).
\end{equation}
Presumably the limit $k \ll 1$ corresponds to when the area of the string worldsheet is large compared to $\Delta$, $h \gg \Delta$. Thus in this limit $ k^2 \ll 1$ and $ h^2 \gg 1$ because of which we can drop the $\mc{O}(k^2/h^2)$ corrections. The final area corrected Nambu-Goto action becomes:
\begin{equation} \label{eqn:modified-NG-action}
	S'_{NG} = -T \int \diff ^2 x \left( \sqrt{- h} - \frac{1}{2} \frac{k \Delta}{\sqrt{- h}} \right).
\end{equation}
This action has, to the best of our knowledge, not been explored in the string theory literature. However, such an inverse determinant term arises very naturally in the LQG approach and goes by the name of ``inverse triad'' corrections, in the context of the 3D volume of spatial hypersurfaces.

@@

We have presented in this work a possible way to formulate string theory in background independent manner.

@@

\begin{itemize}
	\item We adopt the central result of LQG which is quantization of area
	\item We seek to describe the behavior of a string in an arbitrarily curved background geometry.
	\item Connection variables are used in LQG to describe geometry. So we should search for non-metrical description of coupling string geometry to bulk.
	\item For arbitrarily curved bulk geometry, global Poincare symmetry of embedding fields is meaningless.
	\item Elevate global Poincare to local Poincare following the lessons of gauge theory, which leads to introduction of extra "holographic" co-ordinate.
	\item Investigate physical implications of holographic co-ordinate for string phenomenology in this picture.
\end{itemize}

>>>>

\author{Deepak Vaid}
\email{dvaid79@gmail.com}
\affiliation{}
\author{Luigi Teixeira de Sousa}
\email{luigi.tiraque@gmail.com}
\affiliation{Universidade Federal de São Carlos, Brasil}

@@

	The second is an action in which the background geometry is described in terms of the pullback of the connection which describes the bulk geometry to the worldsheet. The resulting action is in the form of a gauged sigma model, where the spacetime co-ordinates are now vectors which transform under $ISO(D,1)$. We find that for the particular case of a constant background connection the action reduces to the bimetric action discussed above. We discuss classical solutions and quantization strategies for this action and its implications for the broader program of unifying string theory and loop quantum gravity.
\end{abstract}

@@

In a very real sense string theory is a quantum theory of geometry. After all, the Nambu-Goto action:
\begin{equation}\label{eqn:ng-action}
	S_{NG} = -T \int \diff ^2 x \sqrt{- \det h}
\end{equation}
is nothing more than the \emph{area} of the string worldsheet. So when we are constructing the quantum theory of the string we are constructing a theory in which the fundamental excitations are geometric in nature. What we would like to do is to modify the string action in a way which incorporates the insight from LQG that at the Planck scale there is a minimum quantum of area. One simple way in which to do this is to modify the action \eqref{eqn:ng-action} as follows:
\begin{equation} \label{eqn:NG-area-corrected}
	S_{M N G} = - T \int \diff ^2 x \sqrt{- (h + k \Delta)}.
\end{equation}
Here, $k = l_{pl} / l_s$, where $l_{pl}$ is the Planck scale, the scale at which quantum geometric effects become important; $l_s$ is the string scale where quantum geometric effects due to area quantization become small; $\Delta$ is a scalar density of weight 2 which can be taken to be the determinant of an antisymmetric worldsheet tensor $\Delta _{a b} (x)$ and corresponds to the square of the quantum of area at the point $x$ where the edges of the bulk spin network state pierce the WS and endow it with quanta of area as described in the previous section. In the limit that $l_s \gg l_{pl}$, $k \ll 1$ and the string action reduces to the conventional Nambu-Goto action with small corrections coming from quantum geometry. In this limit we can expand the integrand as follows:
\begin{equation} \label{eqn:inverse-metric-correction}
	\sqrt{- (h + k \Delta)} \sim \sqrt{- h}\left(1 - \frac{1}{2} \frac{k \Delta}{(- h)} + \mc{O}(k^2/h^2)\right).
\end{equation}
Presumably the limit $k \ll 1$ corresponds to when the area of the string worldsheet is large compared to $\Delta$, $h \gg \Delta$. Thus in this limit $ k^2 \ll 1$ and $ h^2 \gg 1$ because of which we can drop the $\mc{O}(k^2/h^2)$ corrections. The final area corrected Nambu-Goto action becomes:
\begin{equation} \label{eqn:modified-NG-action}
	S'_{NG} = -T \int \diff ^2 x \left( \sqrt{- h} - \frac{1}{2} \frac{k \Delta}{\sqrt{- h}} \right).
\end{equation}
This action has, to the best of our knowledge, not been explored in the string theory literature. However, such an inverse determinant term arises very naturally in the LQG approach and goes by the name of ``inverse triad'' corrections, in the context of the 3D volume of spatial hypersurfaces.

@@

In this article we have presented a phenomenological modification of the Nambu-Goto action, inspired on the one hand, by the central result of loop quantum gravity, according to which area and other geometric observables are quantized at the Planck scale. In particular the spectrum of the area operator is discrete and the smallest eigenvalue is non-zero (of the order $l_{pl}^2 \sim 10^{-68}\,\text{m}^2$). On the other hand we also took as a working hypothesis the statement that string theory, or some suitable version of it, provides the correct description of the low energy physics of the world around us. We then searched for the way in which both of these paradigms can be incorporated into a single theoretical model. The form of the action shown in \eqref{eqn:NG-area-corrected} appears, at least superficially

Let us recall the steps which have led us here. Our original motivation was to continue to expand on the work which was started by one of us several years ago \cite{Vaid2017Connecting} and ask whether it is possible to construct an alternative to the usual action for the bosonic string in which the 2d geometry of the string worldsheet is coupled to the bulk geometry, not through the bulk metric $G_{\mu\nu}$, but through the pullback of a bulk connection to the worldsheet. This allows us to define a covariant derivative on the worldsheet which acts on the embedding co-ordinate fields $X^\mu$

@@

\begin{itemize}
	\item We adopt the central result of LQG which is quantization of area. 
	\item We seek to describe the behavior of a string in an arbitrarily curved background geometry.
	\item Connection variables are used in LQG to describe geometry. So we should search for non-metrical description of coupling string geometry to bulk.
	\item For arbitrarily curved bulk geometry, global Poincare symmetry of embedding fields is meaningless.
	\item Elevate global Poincare to local Poincare following the lessons of gauge theory, which leads to introduction of extra "holographic" co-ordinate.
	\item Investigate physical implications of holographic co-ordinate for string phenomenology in this picture.
\end{itemize}
